---
title: "R Basics"
author: "Introduction to R (online course)"
output:
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Agenda

Basic **operators** and **data structures** in R:

- Operators
- Basic data types and structures
- Data frames.


## Operators

Types:

- **Assignment** operators: `<-`
- **Arithmetic** operators: `+-*/^`
- **Relational** operators: `<>==`
- **Logical** operators: `&|!`


## Assignment  {.columns-2}

Set variables `x`, `y`, `z`:

```{r}
x <- 1
x

y <- "Hello, R!"
z <- TRUE
```

<br/>

Get variables list:

```{r}
ls() # get vars list
rm(z) # remove z var
ls() # check result
```


## Arithmetic and math functions {.columns-2}

- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Exponentiation: `^`
- Division: `/`
- Modulo: `%%`
- Integer division: `%/%`.

_

<br />

Arithmetic operations:
```{r}
(1 / 10 * (100 - 1000)) ^ 2
```

Math functions:
```{r}
abs(log(cos(pi/exp(1))))
```


## **Relational** and **Logical** operators

Example:
```{r}
(1 >= 10) | ((3.14 == 3.14) & !FALSE)
```

<font size="5">
See also [R Operators](https://www.datamentor.io/r-programming/operator/).
</font>



## Data types

Data types:

- **scalars**:  numeric, character or logical values
- **vectors** (one dimensional array): numeric, character or logical value; all elements have the same data type.
- **matrices** (two dimensional array): numeric, character or logical values; all elements have the same data type.
- **lists**: the different items on that list most likely differ in length, characteristic, and type of activity that has to be done. 
- **data frames** (two-dimensional objects): numeric, character or logical values; within a column all elements have the same data type, but different columns can be of different data type.



## Data types (practice)

Vector and list:
```{r}
c(x, 2, 3)

list(a = y, b = c(x, 2, 3), с = NA)
```


##

Matrix and array:
```{r}
matrix(data = 1:15, nrow = 3, ncol = 5)

array(data = 1:24, dim = c(3, 4, 2))
```


##

Data frame:
```{r}
data.frame(
  col1 = 1:10,
  col2 = seq(1, 20, by = 2),
  col3 = rep(Sys.time(), times = 10)
)
```


## Conclusion

<font size="4">
[Back to Course program](/StarRter/)
</font>

