---
title: "R Basics"
author: "Dmitry Petukhov"
date: "`r Sys.Date()`"
output:
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Agenda

Basic operators and data structures in R:

- Arithmetic and operators
- Basic data types and structures
- Data frames.



## Arithmetic
Arithmetic:

- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^`
- Modulo: `%%`.


## Arithmetic (practice)

```{r}
(1 / 10 * (100 - 1000)) ^ 2
```

```{r}
abs(log(cos(pi/exp(1))))
```


## Operators

Operators types:

- Arithmetic operators
- Relational operators
- Logical operators
- Assignment operators.

```{r}
(1 >= 10) | ((3.14 == 3.14) & !FALSE)
```
<font size="4">
See also [R Operators](https://www.datamentor.io/r-programming/operator/).
</font>


## Assignment

Set `x`, `y`, `z`:

```{r}
x <- 1
x

y <- "Hello, R!"
z <- TRUE
```

## Assignment (2)

Get variables list:

```{r}
ls() # get vars list
rm(z) # remove z var
ls() # check result
```


## Data types

Data types:

- **scalars**
- **vectors** (one dimensional array): can hold numeric, character or logical values. The elements in a vector all have the same data type.
- **matrices** (two dimensional array): can hold numeric, character or logical values. The elements in a matrix all have the same data type.
- **lists**: the different items on that list most likely differ in length, characteristic, and type of activity that has to be done. 
- **data frames** (two-dimensional objects): can hold numeric, character or logical values. Within a column all elements have the same data type, but different columns can be of different data type.



## Data types (practice)

Vector and list:
```{r}
c(x, 2, 3)

list(a = y, b = c(x, 2, 3), Ñ = NA)
```


##

Matrix and array:
```{r}
matrix(data = 1:15, nrow = 3, ncol = 5)

array(data = 1:24, dim = c(3, 4, 2))
```


##

Data frame:
```{r}
data.frame(
  col1 = 1:10,
  col2 = seq(1, 20, by = 2),
  col3 = rep(Sys.time(), times = 10)
)
```


