---
title: "Data Frames and Exploratory Data Analysis"
author: "Dmitry Petukhov"
output: 
  html_document:
    df_print: paged
    code_folding: hide
    keep_md: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: flatly
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set_env_opt, include=FALSE}
options(max.print = 1e3, scipen = 999, width = 1e2)
options(stringsAsFactors = F)
```



## Install packages

Install dataset package:
```{r}
install.packages("datasets") 

# invoke interactive help (warn: it's worked only in IDE)
?datasets # short version
library(help = "datasets") # or complete list of datasets
```

Loading / attaching package:
```{r}
library(datasets)
```



## Create data frame

Create new data frame:
```{r}
df <- data.frame(
  col1 = 1:10,
  col2 = seq(1, 20, by = 2),
  col3 = rep(Sys.time(), times = 10)
)

df
```

...or load existing data frame:

```{r}
library(datasets)
library(dplyr)

# load
stocks <- EuStockMarkets 

# transform time-series to data table
stocks <- as_tibble(stocks)
stocks
```



## Discovery data frame structure

View data:
```{r}
# You can code in this style
as_tibble(head(stocks))

# ...or this
stocks %>% tail %>% as_tibble
```


Data slices:
```{r}
stocks[1, ] %>% as_tibble

stocks[10:20, ] %>% as_tibble

stocks[10:20, c(1, 3)] %>% as_tibble
```


Data frame metadata:
```{r}
print(dim(stocks))

print(str(stocks))

print(summary(stocks))

skimr::skim_to_list(stocks)

```

## Descriptive statistics in R

Statistics types:

- Center: `mean(), median()`
- Spread: `sd(), var(), IQR(), mad()`
- Range: `min(), max(), quantile()`
- Position: `first(), last(), nth()`
- Count: `n(), n_distinct()`



```{r}
min(stocks$DAX)
mean(stocks$CAC)
max(stocks$FTSE)
```


```{r}
library(tidyr)

funs <- list(min, mean, median, sd, var, IQR, max)

stocks %>% 
  gather(stock, close) %>% 
  group_by(stock) %>% 
  summarise_all(funs) %>% 
  as_tibble
```

See also: https://www.statsandr.com/blog/descriptive-statistics-in-r/ 


## Vizualization
```{r}
library(ggplot2)

df <- stocks %>% gather(stock, close)
```

### Histogram

```{r}
hist(stocks$DAX)
```



### Boxplot

```{r}
boxplot(df$close ~ df$stock)

boxplot(log(df$close) ~ df$stock)
```



### Scatterplot
```{r}
plot(stocks$DAX, stocks$SMI)
```


### Correlation plot
```{r}
library(corrplot)

fake_trades <- stocks$DAX - rnorm(nrow(stocks), mean = 1000, sd = 2000)
              
M <- cor(stocks %>% mutate(FAKE = fake_trades))
print(M)

corrplot(M)

```

```{r}
corrplot(M,
         method = "square",
         order = "hclust", diag = F, tl.col = "black", tl.cex = 0.7,
         title = "Stocks correlation matrix",
         mar = c(0,1,2,0))
```


### Density plot

```{r}
plot(density(stocks$DAX))

plot(density(na.omit(stocks$DAX - lag(stocks$DAX))))
```





## Conclusion

<font size="4">
[Back to Course program](/StarRter/)
</font>



