---
title: "Importing and Exporting Data"
author: "Dmitry Petukhov"
date: "`r Sys.Date()`"
output: 
  html_document:
    df_print: paged
    code_folding: hide
    keep_md: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: flatly
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set_env_opt, include=FALSE}
options(max.print = 1e3, scipen = 999, width = 1e2)
options(stringsAsFactors = F)
```


## Formats and data sources

Available formats:

- **CSV files**: various separators, various encoding, plain or compressed (.gz, .bz2)
- **Excel**
- **Database**: PostgreSQL, SQL Server, Oracle DB 
- **Big Data clusters**: Spark, Hive, Impala
- **Web**
- RDS, RData
- SPSS, SAS, etc.


From sources:

- **local disc**
- **http(s)://, file://, ftp://**
- **Cloud storage services**, such as Azure Blob Storage.



## Import data

### Import data from `CSV`

Import from local storage:
```{r}
library(data.table)

products_csv <- fread("data/products.csv")
products_csv
```

Import from web (this repo in github):
```{r}
products_web <- fread("https://raw.githubusercontent.com/codez0mb1e/StarRter/master/data/products.csv")
products_web
```

Compare results:
```{r}
dim(products_web)
names(products_web)

stopifnot(
  dim(products_web) == dim(products_csv),
  names(products_web) == names(products_csv)
)
```



### Import data from `Excel`

```{r}
library(readxl)

# read workbook sheets
excel_sheets(path = "data/products.xlsx")

# import sheet of interest
products_xlsx <- read_excel(path = "data/products.xlsx", sheet = "products data")
products_xlsx
```



### Import from `SQL Server`

```
library(DBI)
library(odbc)

# NOTE: change on actual connection string
conn <- dbConnect(odbc(), 
                  .connection_string = "Driver={SQL Server};Driver={SQL Server};Server=<server_db>;database=<db_name>;Trusted_Connection=yes;", 
                  timeout = 10)

products_sql <- dbGetQuery(conn, "select * from dbo.products")
```


## Export data to `CSV`

### Export to CSV

Filter data for export:
```{r}
library(dplyr)

products_csv %>% 
  count(department, brand_ty, sort = T)
 
new_products <- products_csv %>% 
  filter(department == "PHARMA" & brand_ty == "PRIVATE") %>% 
  mutate_if(is.character, tolower) %>% 
  select(-x5)

new_products
```

Export:
```{r}
write.table(new_products, "data/new_products.csv", 
            sep = ",", 
            row.names = F)
```

And now check result via 'Import dataset' wizard in RStudio.


## Conclusion

<font size="4">
[Back to Course program](/StarRter/)
</font>


